Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-04-13 main.Rmd
aperiodic section update - diagram and caption

Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.<!-- $$G'= Alter^{2}_{(1,-1)}(BG, Group_3(B))$$ and --><!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules. Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$. Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$. Hence using \ref:{def:quasicircular}, then we will have:

@@

\begin{definition}\label{aperiodic}
An {\bf aperiodic cyclic granularity} $A_{B, M}$ relates aperiodic linear granularity $M$ and bottom granularity $B$, if
\begin{equation}\label{eq:aperiodic}
\begin{split}
A_{B, M}(z) & =
\begin{cases}
i, & \text{for}\quad z \in T_{i_j} \\
0  & \text{otherwise} 
\end{cases}
\end{split}
\end{equation}
where,
$z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
$T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$.
\end{definition}

@@

**Example**: \autoref{fig:aperiodic-example} shows two aperiodic linear granularity $M_1$ and $M_2$. <!--Each granule is represented by a box. The diagram also illustrates that the granules that overlap share elements from the underlying time domain. -->The first slot in the diagram shows the index set and $A(B, M)$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.

**Discussion**: Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labor day and M can be conceived as the aperiodic linear granularity denoting any holidays. <!--The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.-->

@@

```
<!-- #  -->
<!-- # <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. --> -->

>>>>

Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules.<!-- $$G'= Alter^{2}_{(1,-1)}(BG, Group_3(B))$$ and --><!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules. Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$. Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$. Hence using \ref{def:quasicircular}, then we will have:

@@

\begin{definition}\label{def:aperiodic}
An {\bf aperiodic cyclic granularity} $A_{B, M}$ relates a linear granularity $M$ to the bottom granularity $B$, if
\begin{equation}\label{eq:aperiodic}
\begin{split}
A_{B, M}(z) & =
\begin{cases}
i, & \text{for}\quad z \in T_{i_j} \\
0  & \text{otherwise} 
\end{cases}
\end{split}
\end{equation}
where,
$z \in \mathbb{Z}_{\geq 0}$ denotes the index set,
$B$ denotes a full-integer labelled bottom granularity which groups into linear granularity $M$ but not periodically, $T_{i_j}$ are the sets of indices of $B$ describing aperiodic linear granularities $M_{i}$ such that $M_{i}(j) = \bigcup_{z \in T_{i_j}}B(z)$, and $M = \bigcup_{i=1}^{n}M_{i}$.
\end{definition}

@@

Consider a school semester which lasts for 18.3 weeks, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks after which there is a study break of 1 week before final exams which continue for 1.3 weeks. Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively. Both $M_1$ and $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the semester week type are aperiodic with respect to days ($B$) or weeks ($G$). Hence $A_{B, M}$ denoting day of the semester week type would be an aperiodic cyclic granularity. This is because the placement of the semester within an year would vary across years (\autoref{fig:aperiodic-example} c). \autoref{fig:aperiodic-example} b shows the stack display of time with granules representing same categories (semester break/in-session) stacked on top of one another. \autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}. It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester remains constant over years. Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks.

<!--Each granule is represented by a box. The diagram also illustrates that the granules that overlap share elements from the underlying time domain. The first slot in the diagram shows the index set and $A(B, M)$ consists of three categories, $L(1)$ corresponding to the first aperiodic event $M_1$, $L(2)$ corresponding to aperiodic event $M_2$ and the $L(0)$ corresponding to no event/aperiodic linear granularity.-->

<!-- Here, $M_1$ can be an aperiodic linear granularity denoting block leaves during Christmas, $M_2$ can be the one denoting Easter, $M_n$ can be Labor day and M can be conceived as the aperiodic linear granularity denoting any holidays. <!--The number of categories obtained coincides with the number of aperiodic linear granularities considered and an additional category representing no event.-->

@@

```
<!-- #  -->
<!-- # <!-- To consider the exhaustive set of temporal regularities that might exist in the data, we can also define temporal granularities based on the hierarchical structure of a calendar. -->
